Improved Bounds for Online Balanced Repartitioning

### 2020-03-29 paper/paper.tex
PPL reading
Reading and reviewing

\begin{theorem}
	The competitive ratio of any deterministic online algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$ for any $k\geq 3$.
\end{theorem}

@@

\begin{lemma}	\label{lemma:rebalancecost}
    The cost of repartitioning at Line \ref{line:rebalance} is at most $2\cdot\OPT$.
\end{lemma}
\begin{proof}
    Consider the repartitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
    Let $M \subset V$ denote the set of nodes that migrate during this process.
	Let $M^-$ and $M^+$ denote the subset of nodes that (respectively)
    enter or leave their original cluster during the repartitioning.    
    Then,
    $M = M^+ \cup M^-$.
    Since $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
    the distance before the repartitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.
    Analogously,
     the distance afterwards is $\dist(P_{t},P_I) \geq | M^+|$.
    Thus,
    $|M| \leq \dist(P_{t-1},P_I) + \dist(P_{t},P_I)$.
    By Property \ref{prop:dist<OPT},
    $\dist(P_{t-1},P_I) , \dist(P_{t},P_I) \leq \Delta \leq \OPT$
    and thereby we have	
    $|M| \leq 2\cdot\OPT$.
\end{proof}

\begin{theorem}	\label{thm:upperbound}
    \PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
\end{theorem}
\begin{proof}
      On each inter-cluster request,
     the algorithm enumerates all $\ell$-way partitions of components
     that are in the same (closest) distance of $P_I$.
     That is, 
     once it reaches a configuration $P$ at distance $\Delta = \dist(P, P_I)$,
     it does not move to a configuration
     $P', \dist(P', P_I) > \Delta$,
     before it enumerates all configurations at distance $\Delta$.
     Therefore,
     \PPL eventually reaches $\Delta=\OPT$ and the configuration $P_F$.
%    including the request that completes revealing of all components that are collocated in $P_F$.
    There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{repartition}$
     (i.e., the number of internal edges in $P_F$).
    By Lemma \ref{lemma:rebalancecost},
    each repartition costs at most $2\cdot\OPT$.
    The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$, which implies the competitive ratio.
%    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
%    	So it is 4-competitive (?)}
 \end{proof}

@@

>>>>

\begin{theorem}
  The competitive ratio of any deterministic online algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$ for any $k\geq 3$.
  \maciek{TODO: the exact bound, not asymptotic}
\end{theorem}

@@

\begin{lemma}	\label{lemma:rebalancecost}
    The cost of repartitioning at Line \ref{line:rebalance} is at most $2\cdot\OPT$.
\end{lemma}
\begin{proof}
    Consider the repartitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
    Let $M \subset V$ denote the set of nodes that migrate during this process.
	Let $M^-$ and $M^+$ denote the subset of nodes that (respectively)
    enter or leave their original cluster during the repartitioning.    
    Then,
    $M = M^+ \cup M^-$.
    Since $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
    the distance before the repartitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.
    Analogously,
     the distance afterwards is $\dist(P_{t},P_I) \geq | M^+|$.
    Thus,
    $|M| \leq \dist(P_{t-1},P_I) + \dist(P_{t},P_I)$.
    By Property \ref{prop:dist<OPT},
    $\dist(P_{t-1},P_I) , \dist(P_{t},P_I) \leq \Delta \leq \OPT$
    and thereby we have	
    $|M| \leq 2\cdot\OPT$.
    \maciek{the comma between $\dist$ is confusing. BOTH are bounded?}
\end{proof}

\begin{theorem}	\label{thm:upperbound}
    \PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
    \maciek{Consider not hiding the $2(k-1)\ell$ in the theorem. This is a better bound.}
\end{theorem}
\begin{proof}
      On each inter-cluster request,
     the algorithm enumerates all $\ell$-way partitions of components
     that are in the same (closest) distance of $P_I$.
     That is, 
     once it reaches a configuration $P$ at distance $\Delta = \dist(P, P_I)$,
     it does not move to a configuration
     $P', \dist(P', P_I) > \Delta$,
     before it enumerates all configurations at distance $\Delta$.
     Therefore,
     \PPL eventually reaches $\Delta=\OPT$ and the configuration $P_F$.
%    including the request that completes revealing of all components that are collocated in $P_F$.
    There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{repartition}$
     (i.e., the number of internal edges in $P_F$).
     \maciek{Clarify why the number of internal edges bound the number of repartitions.}
    By Lemma \ref{lemma:rebalancecost},
    each repartition costs at most $2\cdot\OPT$.
    The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$, which implies the competitive ratio.
%    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
%    	So it is 4-competitive (?)}
 \end{proof}

@@

\subsection{Perfect Partition reachable within one swap}

(The version for two clusters.)

Results: LB = UB = k+1.
